---
title: "Ch 3 and 4 Overview"
author: "Ryan Meyer"
date: "4/22/2022"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    theme: flatly
  word_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries 
```{r}
library(tidyverse)
??vegan
?ggplot2
?tidyverse
?filter()
```

# Chapter 3: plotting

The syntax of plotting in ggplot2:
ggplot(data = data, mapping = aes(x = x, y = y, any other aes here)) +
geometries_here()

This can be shortened, I often do this
data %>%
ggplot(aes(x,y)) +
geom_here()


# Starting with some useful geometries 
R has built in data, I'm going to be using the motor trend car data because cars are cool
```{r}
# many different ways to write the code, lots of ways to do this!
mpg
diamonds

ggplot(data = data, mapping = aes(x = x, y = y)) +
  geom_point()

ggplot(data = data, mapping = aes(x = x)) +
  geom_point(mapping = aes(y = y1)) +
  geom_point(aes(y = y2))  

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))


# vs.
mpg %>%
  ggplot(aes(cty, hwy)) +
  geom_point()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point()

# some basic and useful geometries
mpg

mpg %>%
  ggplot(aes(cty, hwy)) +
  geom_point()

mpg %>%
  ggplot(aes(cty, hwy)) +
  geom_jitter()

mpg %>%
  ggplot(aes(cty, hwy)) +
  geom_smooth(se = FALSE)

mpg %>%
  ggplot(aes(x = drv)) +
  geom_bar()

mpg %>%
  ggplot(aes(class)) +
  stat_count()

# can have multiple geometries on the same plot 
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point() +
  geom_smooth(se = TRUE)
  
```

# A useful note
```{r}
# sometimes when you want a continuous variable on your legend it can be useful to set it as a factor

# is this the best example, nah, but it popped into my head pretty quick 
mpg.df <- mpg

mpg %>%
  ggplot(aes(displ, cty, color = hwy)) +
  geom_point()

mpg %>%
  mutate(hwy = as.factor(hwy)) %>%
  ggplot(aes(displ, cty, color = hwy)) +
  geom_point()
```


# Adding Aesthetics to these 
```{r}
# some examples of aesthetics, color, shape, size, alpha 
# again, many different ways to do this, the method used in the textbook is useful for assigning an aesthetic to a particular geometry, putting it in the ggplot statement assigns it to everything, the following code makes the same graph 

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))

mpg %>%
  ggplot(aes(displ, hwy, color = class)) +
  geom_point()

# sometimes you will need to put your aesthetics in the geom statement though, particularly if you want to filter by specific things without changing your data frame

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()

# if you put color in the ggplot statement it does this
mpg %>%
  ggplot(aes(displ, hwy, color = class)) +
  geom_point() +
  geom_smooth()

# the takeaway, there are many ways to do this, one isn't better than the other, but they are useful for particular things
mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth()

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()

```

# Getting a bit more advanced, mutating things within ggplot
```{r}
# you can mutate things within ggplot, you may sometimes want to make a new data frame to make a graph, but you don't need to do this, you can make your mutations within ggplot
mpg_one.df <- mpg.df %>%
  filter(class == "compact")

mpg %>%
  filter(class == "compact") %>%
  ggplot(aes(displ, hwy)) +
  geom_point()

mpg_one.df %>%
  ggplot(aes(displ, hwy)) +
  geom_point()

# but, it also matters where you put these mutations, you can put them within geometries, this is a cool trick!

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)

mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)

mpg %>%
  filter(class == "subcompact") %>%
  ggplot(aes(displ, hwy, color = class)) +
  geom_point() +
  geom_smooth()

mpg %>%
  ggplot(aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```


# Stat Summary, one of the best features of ggplot
## Stat summary will do summary stats for you and plot them it great
```{r}
# mean and standard error plot
mpg %>%
  ggplot(aes(cyl, hwy)) +
  stat_summary(fun = mean, na.rm = TRUE, geom = "point", size = 3) +
  stat_summary(fun.data = mean_se, na.rm = TRUE, geom = "errorbar", width = .3, size = 0.7)

# min max and medium examples
mpg %>%
  ggplot(aes(class, displ)) +
  stat_summary(
    fun.min = min,
    fun.max = max,
    fun = median
  ) +
  theme_isco()

mpg %>%
  ggplot(aes(drv, cty)) +
  geom_boxplot()

mpg %>%
  ggplot(aes(class, displ, group = 1)) +
  stat_summary(fun = mean, na.rm = TRUE, geom = "line", size = 0.5) +
  stat_summary(fun.data = mean_se, na.rm = TRUE, geom = "errorbar", width = 0.3, size = 0.7)
```

# Facet grid and facet wrap - useful for data exploration
```{r}
# facet wrap
mpg %>%
  ggplot(aes(cyl, hwy)) +
  geom_point() +
  facet_wrap(~class)

# facet grid
mpg %>%
  ggplot(aes(cyl, hwy)) +
  geom_point() +
  facet_grid(drv ~ class)
```

# Last thing, making graphs pretty
```{r}
install.packages("ggThemeAssist")
# use that
# can also add a whole host of theme_themes() to make graphs look nice easy
# labs is useful for adding axis labels 
# theme classic is my favorite 
mpg %>%
  ggplot(aes(cyl, cty, color = drv)) +
  stat_summary(fun = mean, na.rm = TRUE, geom = "point", size = 3) +
  stat_summary(fun.data = mean_se, na.rm = TRUE, geom = "errorbar", width = .3, size = 0.7) +
  labs(x = "Cylinders", "City MPG") +
  theme_classic()

# obligatory bar graph - so i made an interesting bar graph...ah jeez
mpg %>%
  ggplot(aes(class)) +
  stat_count(aes(fill = drv)) + theme(axis.line = element_line(linetype = "solid"),
    panel.background = element_rect(fill = NA),
    legend.key = element_rect(fill = NA),
    legend.background = element_rect(fill = NA)) +labs(x = "Class", y = "Count", subtitle = "Number of cars tested by class") #+
  #theme_classic()


mpg %>%
  mutate(class = as.factor(class)) %>%
  mutate(class = fct_reorder(class, desc(val))) %>%
  ggplot(aes(class)) +
  stat_count(aes(fill = drv)) +
  coord_flip() +
  theme_classic()
a <- "aaaaaa"
theme_isco <- function(base_size = 14, base_family = "Times")
{
  theme(
    # PLOT AND LINES
    axis.line = element_line(size = 0.5, linetype = "solid"), 
    axis.ticks = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = NA, linetype = "blank"),
    panel.grid.minor = element_line(colour = NA, linetype = "blank"), 
    panel.background = element_rect(fill = FALSE),
    # LEGEND
    legend.text = element_text(size = 10,  face = "bold"), 
    legend.title = element_text(size = 10, face = "bold"), 
    legend.key = element_rect(fill = NA, size = 0.9), 
    legend.background = element_rect(fill = NA, size = 0.9),
    legend.position = "bottom",
    # AXIS TEXT
    axis.text = element_text(size = 11, face = "bold", colour = "black"), 
    axis.text.x = element_text(size = 11, colour = "black", vjust = .5, hjust = .5, angle = 0), 
    axis.text.y = element_text(size = 11, colour = "black"),
    # AXIS TITLES
    axis.title = element_text(size = 12, face = "bold"),
    #FACETS
    strip.text = element_text(size = 12, face = "bold", colour = "black"),
    strip.background = element_rect(colour="white", fill="white"))
}
```





